Transcriptomic Analysis of Expressions of Biomarkers and Their Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-06-25 Cover_letter_JPM.tex
Updates from Overleaf

% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.
%
% [2021/06/13-17]
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort, CAMK2N1, IL19, and FCGBP, keep ahead of the curve by FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.65$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the "top rank" of the other 17 genes is replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
There is consensus between the TCGA and GSE65858 cohorts that CAMK2N1, IL19, and FCGBP are significant candidates for the biomarker.
%In the consensus of TCGA and GSE65858 cohorts,

%Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf % FDR KM p value of hazards534
% by Validation_GSE65858_survival.R
% Rplot_GSE65858_CoxHR_CAMK2N1_validated.pdf
\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of TCGA HNSCC.}{
%X axis: unadjusted \textit{P} value of Kaplan-Meier survival (-log10 transformed).
%Y axis: multivariate hazard ratio from Cox proportional regression model.
%Dotted line: significant Bonferroni corrected \textit{P} value. 
%\textcolor{red}{Red dots} mark 10 genes (unvalidated), which impact on poor prognosis ($HR>=1.5$). \textcolor{green}{Green dots} mark 10 genes (unvalidated), which affect on better survival ($HR<=0.5$).
    This cohort has been applied for exploration of the candidate biomarkers.
    Total 9416 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05.
    \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19}, \textcolor{red}{FCGBP} and the other 17 genes (marked in \textcolor{black}{black square}) have hazard ratio (HR) $> 1.8$ or $< 0.6$.\\
    Remarks:\\
    \textcolor{red}{Red spots}: $HR > 1.0$;
    \textcolor{green}{Green spots}: $HR < 1.0$;\\
    X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);\\
    Y-axis: HR of Cox proportional hazard regression model\\
    }
    \label{fig:hazards3}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of GSE65858 cohort}{
    % GSE117973 using the same platform GPL10558
    This HNSCC cohort has been applied for filtering of our candidate genes: "CAMK2N1", "IL19" (two probes), and "FCGBP".
    There are 270 HNSCC participants involved in this cohort.
    The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30330.
    Total 534 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05
    (\textcolor{red}{Red spots}: hazard ratio is greater than 1.0;
    \textcolor{green}{Green spots}: hazard ratio is under than 1.0).\\
    The 22 gene symbol listed on the side: hazard ratio $> 1.8$ or $< 0.6$.\\
    (X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);
    Y-axis: the hazard ratio (HR) under Cox proportional hazard regression model)
    }
    \label{fig:hazards534}
\end{figure}

@@

We also modified the last paragraph in the Material and Methods section, and also described "the p-values are affected by the sample size" in the Discussion section.
The paragraphs marked in \textcolor{red}{Red} from page 2 (line 77) to page 3 (line 129) are the part of the manuscript that has been revised. Also, it is shown here in the following paragraph.
\\[0.3cm]

@@

***...the \textit{P} value is affected by the sample size. % TCGA has the most HNSCC sample size, if combine several GSE65858 and with TCGA HNSCC 
An increase of the sample size will decrease overfitting.

@@

[materials and methods]
National Center for Biotechnology Information Reference Sequence (NCBI) 
%RefSeq Release 38 (November 7, 2009) 
Gene Expression Omnibus (GEO)
[Dataset]
The NCBI GEO repositories is a public repository supporting MIAME-compliant data submissions of microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007}: SOFT file (or MINiML) was used to
download RNA-Seq dataset of HNSCC, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event and RFS event and survival time.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the HumanHT-12 v4 Expression BeadChip, which targets on more than 31,000 annotated genes (47,000 probes) derived from the \acrfull{ncbi} Reference Sequence (RefSeq) Release 38 (November 7, 2009).
% an Affymetrix Human Genome U133A microarray (HG-U133A). 
It has been used as a dataset for candidate selection. \\

@@

***
Third, dissection of ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) due to heterogeneity of 20500 RNA-Seq expression data files.
# deal with return(1), return(2) and return(3) errors, try to solve it
32.2\% error01, ZSWIM2 skip from contingencyTCGA(), it is one group issue in Melt() function.
21.05\% errors in  one group in survdiff on KM cutofFinder_func.R.
0.78\%(159 genes) for unknown reason
%# _marginS_
%n_percent_Bonferroni <- ZSWIM2_2$Freq[1]/sum(ZSWIM2_2$Freq) # 0.4593171 -> 0.4581951
%error01_sample <- which(ZSWIM2$X3==1) #  32.2\% error01: "ZSWIM2 skip from contingencyTCGA()"): one group issue in Melt()
%# => *** re-run
%error02_sample <- which(ZSWIM2$X3==2) # 14.2\% error02: There has only one group in survdiff.
%error03_sample <- which(ZSWIM2$X3==3) # 6.85\% error03: There has only one group in survdiff on KM cutofFinder_func.R.
%#=> Test for difference (log-rank test is multiple Chi-square alone survival time) with P-value; or Test for different KM curve between two groups (seperated by PMM1_median 0 vs 1)
% error05_sample <- which(ZSWIM2$X3==5) # 0.78\%(159 genes) error05: for why?

more in discussion
2-3 supplementary figure Rplot_pvaluePlot_NDFIP1.pdf
iscussion about sliding-window pvalueTex to find the more potential candidate (weak or not weak could be double-checked by GSE65858)

>>>>

% problem of it (少 n=30, and RFS only: validation rate 14/100 而且 10% cutoff)
%x Prognoscan (GSE2837) (n=30 less)  RFS only:  10\% cutoff to yield a \textit{P}-value < 0.05.\\
%找出  3 candidates
During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.
%
% [2021/06/13-17]
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort, CAMK2N1, IL19, and FCGBP, keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.65$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the "top rank" of the other 17 genes is replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
There is consensus between the TCGA and GSE65858 cohorts that CAMK2N1, IL19, and FCGBP are significant candidates for the biomarker.
%In the consensus of TCGA and GSE65858 cohorts,

%Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf % FDR KM p value of hazards534
% by Validation_GSE65858_survival.R
% Rplot_GSE65858_CoxHR_CAMK2N1_validated.pdf
\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_TCGA_HNSCC_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of TCGA HNSCC.}{
%X axis: unadjusted \textit{P} value of Kaplan-Meier survival (-log10 transformed).
%Y axis: multivariate hazard ratio from Cox proportional regression model.
%Dotted line: significant Bonferroni corrected \textit{P} value. 
%\textcolor{red}{Red dots} mark 10 genes (unvalidated), which impact on poor prognosis ($HR>=1.5$). \textcolor{green}{Green dots} mark 10 genes (unvalidated), which affect on better survival ($HR<=0.5$).
    This cohort has been applied for exploration of the candidate biomarkers.
    Total 9416 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05.
    \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19}, \textcolor{red}{FCGBP} and the 17 genes (marked in \textcolor{black}{black square}) have hazard ratio (HR) $> 1.8$ or $< 0.6$.
    The 22 genes, listed on the side, have hazard ratio between 0.6 and 1.5.\\
    Remarks:\\
    \textcolor{red}{Red spots}: $HR > 1.0$;
    \textcolor{green}{Green spots}: $HR < 1.0$;\\
    X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value (log10 transformed);\\
    Y-axis: HR of Cox proportional hazard regression model\\
    }
    \label{fig:hazards3}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=13cm]{Rplot_GSE65858_CoxHR_CAMK2N1_top3FDRKM.pdf}
    \bcaption{Volcano plot of genes under survival analyses of GSE65858 cohort}{
    % GSE117973 using the same platform GPL10558
    This HNSCC cohort has been applied for filtering of our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
    Total 534 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05
    (\textcolor{red}{Red spots}: hazard ratio is greater than 1.0);
    (\textcolor{green}{Green spots}: hazard ratio is under than 1.0).\\
    The 22 genes, listed on the side, have hazard ratio $> 1.8$ or $< 0.6$.\\
    (X-axis: Kaplan-Meier survival estimates, with \acrshort{fdr}-adjusted \textit{P} value with log10 transformed);
    (Y-axis: the hazard ratio (HR) under Cox proportional hazard regression model)
    }
    \label{fig:hazards534}
\end{figure}

@@

We modified the paragraphs in the "Material and Methods" section and "Results" section for the analysis of GSE65858.
We also described "the \textit{P} values are affected by the sample size" in the Discussion section.
The paragraphs marked in \textcolor{red}{Red} from page 2 (line 77) to page 3 (line 129) are the part of the manuscript that has been revised. Also, it is shown here in the following paragraph.
\\[0.3cm]

@@

@@

[materials and methods]

%RefSeq Release 38 (November 7, 2009)

%[Dataset]
The Gene Expression Omnibus (GEO) database, founded by National Center for Biotechnology Information Reference Sequence (NCBI), is a public repository supporting MIAME-compliant data submissions of microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007} was used to
download RNA-Seq dataset (in SOFT or MINiML format) of a HNSCC cohort, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event, RFS event and survival time.
There are 270 HNSCC participants involved in this cohort.
%The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30,330.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the platform GPL10558 (Illumina HumanHT-12 v4.0 Expression BeadChip), which targets on more than 30,330 annotated genes (47,000 probes, derived from the \acrfull{ncbi} Reference Sequence, release 38 on November 7, 2009). % (RefSeq) 
GSE65858 has been used as a dataset for selection of our candidates. \\

[Results]
TCGA HNSCC cohort has been applied for exploration of the candidate biomarkers.
Total 9416 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05.
The 10 genes, named as DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, and DKC1, are significantly associated with the poor prognosis of overall survival (OS). 
The other ten genes were over-expressed in the better survival patients (with hazard ratio <=0.5), named as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (previous name as KIAA1683), and NPB.

CAMK2N1, IL19, FCGBP and the other 17 genes have hazard ratio (HR) $> 1.8$ or $< 0.6$.

%During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.
%
% [2021/06/13-17] GSE65858
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with GSE2837 to yield 3 candidates
In the study of GSE65858 cohort, CAMK2N1, IL19, and FCGBP, keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox'HR ($>1.8$ or $<0.65$) (please see Table \ref{tab:table_top3_GSE65858} listed below,
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002 
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
and also see Figure \ref{fig:hazards534}).
However, the "top rank" of the other 17 genes is replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
This HNSCC cohort has been applied for filtering of our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
    Total 534 genes has \acrshort{fdr}-adjusted \textit{P} value less than 0.05
    (\textcolor{red}{Red spots}: hazard ratio is greater than 1.0);
    (\textcolor{green}{Green spots}: hazard ratio is under than 1.0).\\
    The 22 genes, listed on the side, have hazard ratio $> 1.8$ or $< 0.6$.\\
However, these 22 genes have hazard ratio between 0.6 and 1.5 in the study of TCGA HNSCC.
There is consensus between the TCGA and GSE65858 cohorts that CAMK2N1, IL19, and FCGBP are significant candidates for the biomarker.
The consensus between the TCGA and GSE65858 cohorts in Kaplan-Meier survival and Cox's model

[Discussion]
***...the \textit{P} value is affected by the sample size. % TCGA has the most HNSCC sample size, if combine several GSE65858 and with TCGA HNSCC 
An increase of the sample size will decrease overfitting.

@@

% dissection
%ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) 
Third, we have analyzed the error log of the cutoff finding and Cox modeling, 
the results shows that several situations could halt the program running.
It includes:
\begin{outline}
\1 32.2\% event has one group issue in confusion matrix of Chi-square test in Cox regression
\1 21.05\% errors occur by one group issue in log-rank test (survdiff function in R package "survival") in Kaplan-Meier estimate
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow)
\end{outline}
, which could not be detected prior to the program running.
It might be due to heterogeneity of 20500 RNA-Seq expression data.
%# deal with return(1), return(2) and return(3) errors, try to solve it
%32.2\% error01, ZSWIM2 skip from contingencyTCGA(), it is one group issue in Melt() function.
%21.05\% errors in  one group in survdiff on KM cutofFinder_func.R.
%0.78\%(159 genes) for unknown reason
%# _marginS_
%n_percent_Bonferroni <- ZSWIM2_2$Freq[1]/sum(ZSWIM2_2$Freq) # 0.4593171 -> 0.4581951
%error01_sample <- which(ZSWIM2$X3==1) #  32.2\% error01: "ZSWIM2 skip from contingencyTCGA()"): one group issue in Melt()
%# => *** re-run
%error02_sample <- which(ZSWIM2$X3==2) # 14.2\% error02: There has only one group in survdiff.
%error03_sample <- which(ZSWIM2$X3==3) # 6.85\% error03: There has only one group in survdiff on KM cutofFinder_func.R.
%#=> Test for difference (log-rank test is multiple Chi-square alone survival time) with P-value; or Test for different KM curve between two groups (seperated by PMM1_median 0 vs 1)
% error05_sample <- which(ZSWIM2$X3==5) # 0.78\%(159 genes) error05: for why?

more in discussion section:
from Answer 2-3 supplementary figure Rplot_pvaluePlot_NDFIP1.pdf
discussion about sliding-window pvalueTex to find the more potential candidate (weak or not weak could be double-checked by GSE65858)
